---
title: "Underwater photographs - background data"
output: html_notebook
---


```{r}
library(tidyverse)
library(mvabund)
library(lme4)
library(ggplot2)
library(ggthemes)
library(ggpubr)
```


```{r}
photoquads_raw <- read.csv("data/transect_data_clean.csv") 
photoquads_consol_wide_PA <- photoquads_raw

diver_raw <- read.csv("data/diver_data_clean_v2.csv") 
diver_inc_consol <- diver_raw
```


```{r}
diver_inc_consol_sel <- diver_inc_consol %>% select(c(Location, Period, Algae_Encrusting:Coral))

diver_inc_consol_sel$Rep <- paste(diver_inc_consol$RelativePath,diver_inc_consol$File, sep = "_")

diver_inc_consol_sel$Source <- "Photos"

photoquads_consol_sel <- photoquads_consol_wide_PA %>% ungroup() %>% select(c(-Si_Tr, -Loc_Abr, -Transect))

photoquads_consol_sel$Source <- "Surveys"

photoquads_consol_sel <-  rename(photoquads_consol_sel, Rep=Quadrat)


photo_survey_inc_PA_allreps <- full_join(photoquads_consol_sel, diver_inc_consol_sel)

photo_survey_inc_PA_allreps$Loc_Sou_Per <- paste(photo_survey_inc_PA_allreps$Location, photo_survey_inc_PA_allreps$Source, photo_survey_inc_PA_allreps$Period, sep = "_")


```





#Figure 2b: Photoquadrat vs Photograph - Occurence Frequency Mean Difference
```{r}
photo_survey_inc_PA_allreps_sel <- photo_survey_inc_PA_allreps %>% select(Location, Period, Source, c(Algae_Encrusting:Coral))

photo_survey_PA_long <- pivot_longer(photo_survey_inc_PA_allreps_sel, c(Algae_Encrusting:Coral), names_to = 'Taxa', values_to = 'Presence')


photo_survey_mean_SD <- photo_survey_PA_long %>%
  group_by(Location, Period,Source, Taxa) %>%
  summarise(mean=mean(Presence))

photo_survey_mean_SD <- pivot_wider(photo_survey_mean_SD, names_from = Source,
    values_from = mean)


#manual difference calculation

photo_survey_mean_SD$diff <- photo_survey_mean_SD$Photos-photo_survey_mean_SD$Surveys


photo_survey_diff_mean_SD <- photo_survey_mean_SD %>%
  group_by(Location, Taxa) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))


    #reordered by abundance
    
    habitat_order <- photo_survey_mean_SD %>%
  group_by(Taxa) %>%
  summarise(order=mean(Surveys))
    
           photo_survey_diff_mean_SD <- left_join(photo_survey_diff_mean_SD, habitat_order)

       photo_survey_diff_mean_SD %>%
 ggplot(aes(x=reorder(Taxa, order),mean, colour=Taxa,  shape=Location)) + #, group=Period
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
   geom_point(position = position_dodge(0.6), size=2)+
      xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.5, .5)+
geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
  scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"), 
      values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
      theme_classic()+
                   theme(legend.position = "bottom", 
                  legend.key.size = unit(0.2, 'cm'), #change legend key size
               legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=8))+
     theme(axis.title.y = element_blank()) + 
            theme(axis.text = element_text(size = 8)) +
     theme(axis.title.x = element_text(size = 9)) +
       ggtitle(expression(atop("Photoquadrats vs. Photographs", atop(italic("<--- More Frequent in Photoquadrats - More Frequent in Photographs--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
                guides(color = "none")+
              coord_flip()
```
#Figure 2a: Photoquadrat vs Photograph Occurrence frequency linear regression
```{r}

photo_survey_mean_SD$Taxa <- as.character(photo_survey_mean_SD$Taxa)

photo_survey_mean_SD %>%
 ggplot(aes(`Photos`,`Surveys`)) + #, group=Period
         geom_abline(intercept = 0, slope = 1, size = 0.5) +
         geom_smooth(aes(group=1), method='lm')+
                  geom_point(aes( colour= Taxa, shape=Location),size=2)+
      xlab("Photographs Occurence Frequency")+
            ylab("Photoquadrat Occurence Frequency")+
  theme_classic()+
    scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"), 
                       values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
    theme(legend.key.size = unit(0.2, 'cm'), #change legend key size
          legend.position = "bottom",
               legend.title = element_blank(), #change legend title font size
        legend.text = element_text(size=8))+
                   theme(axis.text = element_text(size = 8)) +
     theme(axis.title = element_text(size = 9)) +
   guides(shape = "none")


```


#Figure 3a - Close (<30cm) Photos vs. and wide Photos (>30cm) - Occurence Frequency Mean Difference
   
    
```{r}
diver_inc_consol <-diver_inc_consol %>% filter(Background == "Yes/Good" | Background == "Ok/Partial") 

diver_inc_consol <- filter(diver_inc_consol, Algae_Encrusting != 0 | Algae_Other != 0| Algae_Kelp != 0| Algae_Turf != 0| Ascidian != 0|Sponge !=0 |  Coral!=0  )

diver_inc_consol_type <- diver_inc_consol %>% select(c(Location, Period, Type, c(Algae_Encrusting:Algae_Other), c(Sponge:Coral)))



diver_inc_consol_type$Type <- gsub("\\~", "", diver_inc_consol_type$Type)
diver_inc_consol_type$Type <- gsub("\\(", "", diver_inc_consol_type$Type)
diver_inc_consol_type$Type <- gsub("\\)", "", diver_inc_consol_type$Type)


diver_inc_consol_type$Type <- gsub("Supermacro  <3cm", "Macro", diver_inc_consol_type$Type)
diver_inc_consol_type$Type <- gsub("Macro 3cm-30cm", "Macro", diver_inc_consol_type$Type)
diver_inc_consol_type$Type <- gsub("Mid 0.3-1m", "Wide-angle", diver_inc_consol_type$Type)
diver_inc_consol_type$Type <- gsub("Wide-angle >1m", "Wide-angle", diver_inc_consol_type$Type)

diver_inc_consol_type_long <- pivot_longer(diver_inc_consol_type, c(Algae_Encrusting:Coral), names_to = 'Taxa', values_to = 'Presence')


diver_inc_type_mean_SD <- diver_inc_consol_type_long %>%
  group_by(Location, Period,Type, Taxa) %>%
  summarise(mean=mean(Presence))

diver_inc_type_mean_SD <- pivot_wider(diver_inc_type_mean_SD, names_from = Type,
    values_from = mean)


#manual difference calculation - Type

diver_inc_type_mean_SD$diff <- diver_inc_type_mean_SD$"Wide-angle"-diver_inc_type_mean_SD$Macro


diver_inc_type_mean_SD <- diver_inc_type_mean_SD %>%
  group_by(Location, Taxa) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))

 diver_inc_type_mean_SD <- left_join(diver_inc_type_mean_SD, habitat_order)

 
  
diver_inc_type_mean_SD %>%
  ggplot(aes(x=reorder(Taxa, order),mean, colour=Taxa, shape=Location)) + 
   geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
   geom_point(position = position_dodge(0.6), size=2)+
   xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.5, .5)+
 geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
         scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"), 
                         values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+

  scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
   theme_classic()+
  theme(legend.position = "bottom", 
                  legend.key.size = unit(0.2, 'cm'), #change legend key size
              legend.title = element_text(size=8), #change legend title font size
       legend.text = element_text(size=8))+
            theme(axis.text = element_text(size = 8)) +
theme(axis.title.y = element_blank()) +
     theme(axis.title.x = element_text(size = 9)) +  
        ggtitle(expression(atop("Close Photos vs. Wide Photos", atop(italic("<--- More Frequent in Close Photos - More Frequent in Wide Photos--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
   guides(colour="none")+
  coord_flip()
```


#Supp Figure S4a - Photoquadrats vs. Close (<30cm) Photos - Occurence Frequency Mean Difference

```{r}
diver_inc_consol_type_Macro <- filter (diver_inc_consol_type, Type=="Macro")

diver_inc_consol_type_Macro$Source <- "Photos"

photo_survey_inc_PA_Macro <- full_join(photoquads_consol_sel, diver_inc_consol_type_Macro)

photo_survey_inc_PA_Macro$Loc_Sou_Per <- paste(photo_survey_inc_PA_Macro$Location, photo_survey_inc_PA_Macro$Source, photo_survey_inc_PA_Macro$Period, sep = "_")

photo_survey_inc_PA_Macro_Long <-  photo_survey_inc_PA_Macro %>% pivot_longer(c(Algae_Encrusting:Coral),names_to = 'Taxa', values_to = 'Presence')


photo_survey_Macro_mean_SD <- photo_survey_inc_PA_Macro_Long %>%
  group_by(Location, Period,Source, Taxa) %>%
  summarise(mean=mean(Presence))

photo_survey_Macro_mean_SD <- pivot_wider(photo_survey_Macro_mean_SD, names_from = Source,     values_from = mean)


photo_survey_Macro_mean_SD$diff <- photo_survey_Macro_mean_SD$Photos-photo_survey_Macro_mean_SD$Surveys


photo_survey_Macro_mean_SD <- photo_survey_Macro_mean_SD %>%
  group_by(Location, Taxa) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))

photo_survey_Macro_mean_SD <- left_join(photo_survey_Macro_mean_SD, habitat_order)



photo_survey_Macro_mean_SD %>%
  ggplot(aes(x=reorder(Taxa, order),mean, colour=Taxa,shape=Location)) + 
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_point(position = position_dodge(0.6), size=2)+
   xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.47, .47)+
geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"),     values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
  scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
   theme_classic()+
  theme(legend.position ="none")+
            theme(axis.text = element_text(size = 8)) +
  theme(axis.title.y = element_blank()) +
     theme(axis.title.x = element_text(size = 9)) +  
        ggtitle(expression(atop("Photoquadrats vs. Close Photographs", atop(italic("<--- More Frequent in Photoquadrats - More Frequent in Close Photos--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
    guides(colour="none")+
  coord_flip()
```
 
#Supp Figure S4b - Photoquadrats vs. Wide (>30cm) Photos - Occurence Frequency Mean Difference

```{r}
diver_inc_consol_type_Wide <- filter (diver_inc_consol_type, Type=="Wide-angle")

diver_inc_consol_type_Wide$Source <- "Photos"

photo_survey_inc_PA_Wide <- full_join(photoquads_consol_sel, diver_inc_consol_type_Wide)

photo_survey_inc_PA_Wide$Loc_Sou_Per <- paste(photo_survey_inc_PA_Wide$Location, photo_survey_inc_PA_Wide$Source, photo_survey_inc_PA_Wide$Period, sep = "_")

photo_survey_inc_PA_WideAng_Long <-  photo_survey_inc_PA_Wide %>% pivot_longer(c(Algae_Encrusting:Coral),names_to = 'Taxa', values_to = 'Presence')


photo_survey_WideAng_mean_SD <- photo_survey_inc_PA_WideAng_Long %>%
  group_by(Location, Period,Source, Taxa) %>%
  summarise(mean=mean(Presence))

photo_survey_WideAng_mean_SD <- pivot_wider(photo_survey_WideAng_mean_SD, names_from = Source,
    values_from = mean)


#manual difference calculation

photo_survey_WideAng_mean_SD$diff <- photo_survey_WideAng_mean_SD$Photos-photo_survey_WideAng_mean_SD$Surveys


photo_survey_WideAng_mean_SD <- photo_survey_WideAng_mean_SD %>%
  group_by(Location, Taxa) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))

photo_survey_WideAng_mean_SD <- left_join(photo_survey_WideAng_mean_SD, habitat_order)


photo_survey_WideAng_mean_SD %>%
  ggplot(aes(x=reorder(Taxa, order),mean, colour=Taxa, shape=Location)) + #, group=Period
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_point(position = position_dodge(0.6), size=2)+
   xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.47, .47)+
 geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"),     values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
  scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
   theme_classic()+
  theme(legend.position = "none", 
                  legend.key.size = unit(0.2, 'cm'), #change legend key size
               legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=8))+
  theme(axis.title.y = element_blank()) +
            theme(axis.text = element_text(size = 8)) +
     theme(axis.title.x = element_text(size = 9)) +  
        ggtitle(expression(atop("Photoquadrats vs. Wide Photographs", atop(italic("<--- More Frequent in Photoquadrats - More Frequent in Wide Photos--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
  coord_flip()

```


#Habitat by Taxa 
```{r}
diver_inc_mat <-  select(diver_inc_consol, Diver,  RelativePath, File,  Location, Date, Date2, Period, Type, Duplicate, Comment, Background, Unknown_Sub, Reason_Unk, Fish_Sub, Inverts_Sub, All_Sub, Sand, Algae_Encrusting, Algae_Turf,Algae_Kelp,   Algae_Other, Sponge, Ascidian, Coral) 

diver_inc_mat_ex0 <-diver_inc_consol %>% filter(Background == "Yes/Good" | Background == "Ok/Partial") 

diver_inc_mat_ex0 <- filter(diver_inc_mat_ex0, Algae_Encrusting != 0 | Algae_Other != 0| Algae_Kelp != 0| Algae_Turf != 0| Ascidian != 0|Sponge !=0 |  Coral!=0  )

diver_inc_mat_tmp <- separate(
  diver_inc_mat_ex0,
  All_Sub,
  into =c("Sub_1", "Sub_2"),
  sep = ",",
   fill = "right",
  )

diver_inc_mat_tmp <- diver_inc_mat_tmp %>% filter(is.na(Sub_2))

diver_inc_mat_tmp <- separate(
  diver_inc_mat_tmp,
  Sub_1,
  into =c("Scientific_Name", "Common_Name"),
  sep = " - ",
  fill = "right",
  )

diver_inc_mat_tmp <- separate(
  diver_inc_mat_tmp,
  Scientific_Name,
  into =c("Scientific_Name", "extra names"),
  sep = "/",
   fill = "right",
  )

species <- as.list(unique(diver_inc_mat_tmp$Scientific_Name))


species_DF<-as.data.frame(species)
diver_inc_mat_tmp <- separate(diver_inc_mat_tmp, Scientific_Name, into = c("Genus", "species"), sep = " ")
genus<-dput(as.character(unique(diver_inc_mat_tmp$Genus)))
```


```{r}
WorMS_record_df <-read.csv(file="data/WorMS_records.csv")

diver_inc_mat_WorMS <-left_join (diver_inc_mat_tmp, WorMS_record_df, by='Genus')


#do fish and gastropods rather than chordates and molluscs
diver_inc_mat_WorMS_class <-    filter(diver_inc_mat_WorMS, class=="Actinopteri"| class=="Elasmobranchii"|class=="Gastropoda")
diver_inc_mat_WorMS_class$taxa<- diver_inc_mat_WorMS_class$class
diver_inc_mat_WorMS_class$taxa<- gsub("Actinopteri", "Fish",diver_inc_mat_WorMS_class$taxa)
  diver_inc_mat_WorMS_class$taxa<- gsub("Elasmobranchii", "Fish",diver_inc_mat_WorMS_class$taxa)
   diver_inc_mat_WorMS_class$taxa<- gsub("Gastropoda", "Gastropod",diver_inc_mat_WorMS_class$taxa)
 
   #Need to remove Shelly Beach as only has 56 gastropods spread over the 3 time periods
diver_inc_mat_WorMS_class <-    filter(diver_inc_mat_WorMS_class, Location !="Shelly Beach")
```




#Figure 3b - Gastropod Photos vs. Fish Photos - Occurence Frequency Mean Difference
```{r}
taxa_av_mancalc <- diver_inc_mat_WorMS_class %>%
select (5, 8, 23:29, 57)

taxa_av_mancalc <- taxa_av_mancalc %>%
  pivot_longer(c(3:9), names_to="Habitat", values_to="presence")

taxa_av_mancalc <- taxa_av_mancalc %>%
filter(Location !="Shelly Beach") %>%
  group_by(Location, Period, taxa, Habitat) %>%
  summarise("mean" = mean(presence)) #, "SE" = sd(presence)/sqrt(n())


taxa_av_mancalc <- pivot_wider(taxa_av_mancalc, names_from = taxa,
    values_from = mean)

# mean of fitted values - across replicate periods


taxa_av_mancalc$diff <- taxa_av_mancalc$Fish-taxa_av_mancalc$Gastropod

taxa_av_mancalc <- taxa_av_mancalc %>%
  group_by(Location,  Habitat) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))

habitat_order_2 <- rename(habitat_order, "Habitat"=Taxa)
taxa_av_mancalc <- left_join(taxa_av_mancalc, habitat_order_2)


taxa_av_mancalc %>%
   ggplot(aes(x=reorder(Habitat, order),mean, colour=Habitat, shape=Location)) + #, group=Period
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
geom_point(position = position_dodge(0.6), size=2)+
      xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.4, .43)+
geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"), 
                         values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
  theme_classic()+
                 theme(legend.position = "bottom", 
                 legend.key.size = unit(0.2, 'cm'), #change legend key size
               legend.title = element_text(size=8), #change legend title font size
       legend.text = element_text(size=8))+
             theme(axis.text = element_text(size = 8)) +
  theme(axis.title.y = element_blank()) +
     theme(axis.title.x = element_text(size = 9)) + 
        ggtitle(expression(atop("Gastropod Photos vs. Fish Photos", atop(italic("<--- More Frequent in Gastropod Photos - More Frequent in Fish Photos--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
guides(colour="none")+
  coord_flip()
        
```

#Supp Figure S4c - Photoquadrats vs. Fish Photos - Occurence Frequency Mean Difference
```{r}


diver_inc_mat_WorMS_Fish <-    filter(diver_inc_mat_WorMS, class=="Actinopteri"| class=="Elasmobranchii")

diver_inc_mat_WorMS_Fish$Source <- "Photos"

diver_inc_mat_WorMS_Fish_sel <- diver_inc_mat_WorMS_Fish %>% select(c(Location, Period, Source, Type, Algae_Encrusting:Coral))


photo_survey_inc_Fish <- full_join(photoquads_consol_sel, diver_inc_mat_WorMS_Fish_sel)



Fish_av_mancalc <- photo_survey_inc_Fish %>%
select (2:12)

Fish_av_mancalc <- Fish_av_mancalc %>%
  pivot_longer(c(4:10), names_to="Habitat", values_to="presence")

Fish_av_mancalc <- Fish_av_mancalc %>%
  group_by(Location, Source, Period, Habitat) %>%
  summarise("mean" = mean(presence)) #, "SE" = sd(presence)/sqrt(n())


Fish_av_mancalc <- pivot_wider(Fish_av_mancalc, names_from = Source,
    values_from = mean)


Fish_av_mancalc$diff <- Fish_av_mancalc$Photos-Fish_av_mancalc$Surveys

Fish_av_mancalc <- Fish_av_mancalc %>%
  group_by(Location,  Habitat) %>%
  summarise(mean=mean(diff), SE =sd(diff)/(sqrt(n())))

Fish_av_mancalc <- left_join(Fish_av_mancalc, habitat_order_2)

  
Fish_av_mancalc %>%
    ggplot(aes(x=reorder(Habitat, order),mean, colour=Habitat, shape=Location)) + #, group=Period
 geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
 geom_point(position = position_dodge(0.6), size=2)+
      xlab("Habitat Categories")+
            ylab("Mean Difference")+
  ylim(-.4, .43)+
geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),position = position_dodge(0.6), width=.6)+
scale_colour_manual(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae", "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"),     values = c("#F0655B", "#C18A15","#306B34","#643C04", "#893168","#FFB9B4","#FFA600"))+
  scale_x_discrete(labels=c("Algae_Turf" = "Turf Algae", "Algae_Other" = "Other Algae",
                              "Algae_Encrusting" = "Encrusting Algae", "Algae_Kelp" = "Kelp Algae", "Sponge" = "Sponges", "Ascidian" = "Ascidians", "Coral" = "Corals"))+
  theme_classic()+
                   theme(legend.position = "none", 
                  legend.key.size = unit(0.2, 'cm'), #change legend key size
               legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=8))+
             theme(axis.text = element_text(size = 8)) +
  theme(axis.title.y = element_blank()) +
     theme(axis.title.x = element_text(size = 9)) + 
        ggtitle(expression(atop("Photoquadrats vs. Fish Photos", atop(italic("<--- More Frequent in Photoquadrats - More Frequent in Fish Photos--->"), "")))) +
       theme(plot.title=element_text(size=15, hjust=0.5, face="bold",  vjust=-1))+
      coord_flip()

```

#Figure 4: number of photographs -  bootstrapping error

```{r}
 diver_inc_consol_5grps <- diver_inc_consol_sel


diver_inc_consol_5grps$randgrp <-  sample(5, size = nrow(diver_inc_consol_5grps), replace = TRUE)


test_manyglm_fit_frm <- expand_grid(Location=c("Bare Is", "Shelly Beach", "Kurnell")) #outside function - run once only...

sample_sizes <- seq(10,180, by=10)
reps <- c(1:5)

diver_inc_boot <- function(n_samp, rep, data) {
test  <-  data %>% #
    group_by(Location)%>%
 filter(randgrp == rep)%>% 
  slice_sample(n=n_samp) 

test_mvabund <-  mvabund(test[3:9])

test_manyglm <- manyglm(test_mvabund~Location, data=test, family=binomial("cloglog"))

test_manyglm_fit_tmp<-predict.manyglm(test_manyglm, newdata=test_manyglm_fit_frm, type="response", se.fit=TRUE)

test_manyglm_fit <- cbind(test_manyglm_fit_frm, test_manyglm_fit_tmp)

rm(test_manyglm_fit_tmp)

test_manyglm_fit

} 



boot_fit <- data.frame()
test_manyglm_fit<- data.frame()

for (rep in reps)for (samp in sample_sizes) {
sample <- diver_inc_boot(samp, rep,    diver_inc_consol_5grps)
  
  sample$samps <- samp
 
  boot_fit <-  rbind(boot_fit,sample )
}


boot_fit_long <- pivot_longer(boot_fit[c(1:8,16)], cols= c(fit.Algae_Encrusting:fit.Coral), names_to = "Taxa", values_to = "Fit")

boot_fit_long$Taxa <- gsub("fit.", "",boot_fit_long$Taxa) 



boot_SE_long <- rename(boot_fit,"SE.Algae_Encrusting"="se.fit.1", "SE.Algae_Turf"="se.fit.2", "SE.Algae_Kelp"="se.fit.3", "SE.Algae_Other"="se.fit.4" , "SE.Sponge"="se.fit.5", "SE.Ascidian"="se.fit.6", "SE.Coral"="se.fit.7")

boot_SE_long <- pivot_longer(boot_SE_long[c(1, 9:16)], cols= c(SE.Algae_Encrusting:SE.Coral), names_to = "Taxa", values_to = "SE")

boot_SE_long$Taxa <- gsub("SE.", "",boot_SE_long$Taxa) 

fit_SE_av <-boot_SE_long %>% group_by(Location, samps, Taxa) %>%
             summarise(Av_error = mean(SE, na.rm = TRUE), SE = sd(SE)/sqrt(5))
 
fit_SE_av$Taxa <- gsub( "Algae_Turf", "a) Turf Algae",fit_SE_av$Taxa )
fit_SE_av$Taxa <- gsub( "Algae_Encrusting", "b) Encrusting Algae",fit_SE_av$Taxa )
fit_SE_av$Taxa <- gsub( "Algae_Other", "c) Other Algae",fit_SE_av$Taxa )
fit_SE_av$Taxa <- gsub( "Algae_Kelp", "d) Kelp Algae",fit_SE_av$Taxa )

fit_SE_av$Taxa <- gsub( "Sponge", "e) Sponges",fit_SE_av$Taxa )
fit_SE_av$Taxa <- gsub( "Ascidian", "f) Ascidians",fit_SE_av$Taxa )
fit_SE_av$Taxa <- gsub( "Coral", "g) Corals",fit_SE_av$Taxa )

#facet SE by taxa
ggplot(fit_SE_av, aes(samps,Av_error,color=Location, shape=Location)) +
  geom_point(size=2)+
       geom_line(stat = 'identity')+
  geom_errorbar(aes(ymin=Av_error-SE, ymax=Av_error+SE), width=.1) +
    facet_wrap(~factor(Taxa,levels=c('a) Turf Algae','b) Encrusting Algae','c) Other Algae','d) Kelp Algae', 'e) Sponges','f) Ascidians','g) Corals')), ncol = 2, scales="free")+ 
        scale_colour_manual(values=c("#EF6461", "#3C7A89", "#966D81"))+ 
  # theme_classic()+ 
   theme_tufte(base_family = "") + 
  theme(axis.line=element_line()) + 
  theme(strip.text = element_text(face="bold", hjust=0.03, margin=margin(l=0, b=1))) + 
scale_x_continuous(name="Number of Photographs", breaks=seq(0,250, by=50))+
  theme(legend.position = c(0.75, 0.1), 
                  legend.key.size = unit(0.4, 'cm'), #change legend key size
               legend.title = element_text(size=9), #change legend title font size
        legend.text = element_text(size=9))+
          theme(axis.text = element_text(size = 8)) +
     theme(axis.title = element_text(size = 9)) + 
  theme(panel.grid = element_line(colour = "NA"))+
    ylab("Average Error of Model Runs")

```


#Supp Figure S5: Number of divers - Bootstrapping error

## #'d out as slow to run. Remove #'s in front of ''' to run

#```{r}

  # install.packages('RcppAlgos')  
     library(RcppAlgos)


#diver combinations function

       combs_function <- function(n_divers, site){
      
  rand30_combs_test <-    as.data.frame(comboSample(unique(site$Diver), m=n_divers, n=length(unique(site$Diver)), seed=42, nThreads = 4))
  rand30_combs_test
                    
       }
       
   #average habitat function
       
       hab_av_func <- function(n_samps, site){
comb_tmp <- as.data.frame(all_combs_long[,n_samps]) #need 1 to iterate through all combinations - n_combs
colnames(comb_tmp)[1] <- 'Diver'

hab_data_samp<-left_join(comb_tmp,site )

comb_tmp <- comb_tmp %>% drop_na()

hab_data_samp$n_divers <- length(unique(comb_tmp$Diver))

hab_data_samp <- hab_data_samp %>% drop_na(Location)

hab_data_sum <- hab_data_samp %>%
  group_by(Location, n_divers)%>%
summarise(Algae_Encrusting = sum(Algae_Encrusting)/n(), Algae_Turf= sum(Algae_Turf)/n(), Algae_Kelp= sum(Algae_Kelp)/n(), Algae_Other= sum(Algae_Other)/n(), Sponge= sum(Sponge)/n(), Ascidian= sum(Ascidian)/n(), Coral= sum(Coral)/n())

hab_data_sum
 
       }
       
      #calculations for bare Island 
       
bare_is <- diver_inc_consol%>%
  filter(Location=="Bare Is") #need to iterate through sites
 
      n_divers_str <- c((length(unique(bare_is$Diver))-1):1)
      all_combs <- as.data.frame( comboSample(unique(bare_is$Diver), m=length(unique(bare_is$Diver)), n=1, seed=42, nThreads = 4))
      
       for (n in n_divers_str) {
comb_tmp <- combs_function(n,    bare_is)
  
all_combs <-bind_rows(all_combs,comb_tmp) 

}

       
       
  all_combs_long <- t(all_combs)     


n_combs <- c(1:ncol(all_combs_long))



hab_data_bare <- data.frame()

for (n in n_combs){
hab_data_sum <-hab_av_func(n, bare_is)

   hab_data_bare <- rbind(hab_data_bare,hab_data_sum)


}
#repeat for shelly beach

shelly <- diver_inc_consol%>%
  filter(Location=="Shelly Beach") #need to iterate through sites
 
      n_divers_str <- c((length(unique(shelly$Diver))-1):1)

            all_combs <- as.data.frame(comboSample(unique(shelly$Diver), m=length(unique(shelly$Diver)), n=1, seed=42, nThreads = 4))

         for (n in n_divers_str) {
comb_tmp <- combs_function(n,    shelly)
  
all_combs <-bind_rows(all_combs,comb_tmp) 

}

      
  all_combs_long <- t(all_combs)     

n_combs <- c(1:ncol(all_combs_long))


hab_data_shelly <- data.frame()

for (n in n_combs){
hab_data_sum <-hab_av_func(n, shelly)

   hab_data_shelly <- rbind(hab_data_shelly,hab_data_sum)

}

#repeat for Kurnell

kurnell <- diver_inc_consol%>%
  filter(Location=="Kurnell") #need to iterate through sites
 
      n_divers_str <- c((length(unique(kurnell$Diver))-1):1)

            all_combs <- as.data.frame(comboSample(unique(kurnell$Diver), m=length(unique(kurnell$Diver)), n=1, seed=42, nThreads = 4))

   
       
       for (n in n_divers_str) {
comb_tmp <- combs_function(n,    kurnell)
  
all_combs <-bind_rows(all_combs,comb_tmp) 

}

      
  all_combs_long <- t(all_combs)     

n_combs <- c(1:ncol(all_combs_long))


hab_data_kurnell <- data.frame()

for (n in n_combs){
hab_data_sum <-hab_av_func(n, kurnell)

   hab_data_kurnell <- rbind(hab_data_kurnell,hab_data_sum)

}

hab_data_n_diver <- rbind(hab_data_bare, hab_data_kurnell, hab_data_shelly)


hab_data_df_long <- hab_data_n_diver
hab_data_df_long$rep <- c(1:nrow(hab_data_df_long))
hab_data_df_long <-  pivot_longer(hab_data_df_long, c(Algae_Encrusting:Coral), names_to = 'Taxa', values_to = 'av_occ')

dodge <- position_dodge(width=0.3)



hab_data_df_long_error <- hab_data_df_long %>% 
  group_by(Location, n_divers, Taxa) %>%
             summarise(SE = sd(av_occ)/sqrt(n()))

hab_data_df_long_error$Taxa <- gsub( "Algae_Turf", "a) Turf Algae",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa <- gsub( "Algae_Encrusting", "b) Encrusting Algae",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa <- gsub( "Algae_Other", "c) Other Algae",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa <- gsub( "Algae_Kelp", "d) Kelp Algae",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa<- gsub( "Sponge", "e) Sponges",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa<- gsub( "Ascidian", "f) Ascidians",hab_data_df_long_error$Taxa )
hab_data_df_long_error$Taxa<- gsub( "Coral", "g) Corals",hab_data_df_long_error$Taxa )


ggplot(hab_data_df_long_error, aes(n_divers,SE,color=Location, shape=Location, group=Location)) +
  geom_point(size=2, position = dodge)+
       geom_line(stat = 'identity')+
 facet_wrap(~factor(Taxa,levels=c('a) Turf Algae','b) Encrusting Algae','c) Other Algae','d) Kelp Algae', 'e) Sponges','f) Ascidians','g) Corals')), ncol = 2, scales="free")+ 
        scale_colour_manual(values=c("#EF6461", "#3C7A89", "#966D81"))+ 
     theme_tufte(base_family = "") + 
  theme(axis.line=element_line()) + 
  theme(strip.text = element_text(face="bold", hjust=0.03, margin=margin(l=0, b=1))) + 
  theme(legend.position = c(0.75, 0.1), 
                  legend.key.size = unit(0.4, 'cm'), #change legend key size
               legend.title = element_text(size=9), #change legend title font size
        legend.text = element_text(size=9))+
          theme(axis.text = element_text(size = 8)) +
     theme(axis.title = element_text(size = 9)) + 
  theme(panel.grid = element_line(colour = "NA"))+
  scale_x_continuous(name="Number of Divers", breaks=seq(0,30, by=10))+
  ylab("Average Error of Model Runs")

#```

##Multivariate abundance tests - #'d out as slow to run. Remove #'s in front of ''' to run

#Photographs vs Photoquadrats mvabund - all data
#```{r}
Pho_Sur_Inc_mvabund <-  mvabund(photo_survey_inc_PA_allreps[5:11])


Pho_Sur_Inc_manyglm_LS <- manyglm(Pho_Sur_Inc_mvabund~Location*Source, data=photo_survey_inc_PA_allreps, family=binomial("cloglog"))
Pho_Sur_Inc_ANOVA_test_LS <- anova(Pho_Sur_Inc_manyglm_LS,  nBoot=1000, test="LR", cor.type = "I", p.uni="adjusted")
Pho_Sur_Inc_ANOVA_test_LS

#```

#Close Photographs vs Wide Photographs mvabund
#```{r}
habitat_Style_mvabund <-  mvabund(diver_inc_consol_type[4:10])


habitat_Style_manyglm <- manyglm(habitat_Style_mvabund~Location*Type, data=diver_inc_consol_type, family=binomial(link = "cloglog"))

### hash out to avoid running  - 30min runtime
habitat_Style_ANOVA <- anova(habitat_Style_manyglm,  nBoot=1000, test="LR", cor.type = "I", pairwise.comp=~diver_inc_consol_type$Location*Type, p.uni="adjusted")

habitat_Style_ANOVA
#```

#Close-up Photographs vs Photoquadrats mvabund
#```{r}
Pho_Sur_Inc_mvabund_Macro <-  mvabund(photo_survey_inc_PA_Macro[5:11])


Pho_Sur_Inc_manyglm_Macro_LS <- manyglm(Pho_Sur_Inc_mvabund_Macro~Location*Source, data=photo_survey_inc_PA_Macro, family=binomial("cloglog"))

Pho_Sur_Inc_ANOVA_test_Macro_LS <- anova(Pho_Sur_Inc_manyglm_Macro_LS,  nBoot=1000, test="LR", cor.type = "I", p.uni="adjusted")

Pho_Sur_Inc_ANOVA_test_Macro_LS

#```

#Wide-angle Photographs vs Photoquadrats mvabund
#```{r}
Pho_Sur_Inc_mvabund_Wide <-  mvabund(photo_survey_inc_PA_Wide[5:11])


Pho_Sur_Inc_manyglm_Wide_LS <- manyglm(Pho_Sur_Inc_mvabund_Wide~Location*Source, data=photo_survey_inc_PA_Wide, family=binomial("cloglog"))

Pho_Sur_Inc_ANOVA_test_Wide_LS  <- anova(Pho_Sur_Inc_manyglm_Wide_LS ,  nBoot=1000, test="LR", cor.type = "I", p.uni="adjusted")

Pho_Sur_Inc_ANOVA_test_Wide_LS 

#```

#Fish photographs vs Gastropod photographs mvabund
#```{r}
habitat_byclass_mvabund <-  mvabund(diver_inc_mat_WorMS_class[22:28])


habitat_byclass_manyglm <- manyglm(habitat_byclass_mvabund~Location*taxa, data=diver_inc_mat_WorMS_class, family=binomial(link = "cloglog"))

habitat_byclass_ANOVA <- anova(habitat_byclass_manyglm,  nBoot=1000, test="LR", cor.type = "I", pairwise.comp=~diver_inc_mat_WorMS_class$Location*taxa, p.uni="adjusted")


habitat_byclass_ANOVA

#```

#Fish photographs vs photoquadrats mvabund
#```{r}
Pho_Sur_Inc_mvabund_Fish <-  mvabund(photo_survey_inc_Fish[5:11])

Pho_Sur_Inc_manyglm_Fish_LS <- manyglm(Pho_Sur_Inc_mvabund_Fish~Location*Source, data=photo_survey_inc_Fish, family=binomial("cloglog"))

Pho_Sur_Inc_ANOVA_Fish <- anova(Pho_Sur_Inc_manyglm_Fish_LS,  nBoot=1000, test="LR", cor.type = "I", p.uni="adjusted")

Pho_Sur_Inc_ANOVA_Fish

#```